Measuring causal influence with\\ back-to-back regression: the linear case - supplementary material

### 2019-09-25 paper/meg_exps.tex
meg correction

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is displayed in \ref{fig:meg_evoked}
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We are here interested in identifying the features of word that modulate (i.e. cause) brain activity. We consider four distinct, but colinear features.
%
"Word Length' refers to the total number of letters. Word Length is expected to specifically cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
'Word Frequency', as derived with WordFreq \citep{speerwordfreq}, indexes how frequently each word appears in Dutch (using the Zipf logarithmic scale of \citep{van2014subtlex}). Word Frequency is expected to specifically cause a variation in the late evoked MEG responses (i.e. from 400 ms), because it variably engages semantic processes in the temporal cortices \citep{kutas2011thirty}.
%
'Word Function' indicates whether each word is a content word (i.e. a noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition, a conjunction, a determinant, a pronoun or a numeral), and was derived from Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has not been thouroughly investigated with MEG. Its causal contribution to reading processes in the brain thus remains unclear.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of "Word Length" and "Word Frequency":
\begin{equation}
  dummy = z(length) + z(frequency) + \mathcal{N}
\end{equation}
where $z$ transforms features into zero-mean and unit variance vectors and $\mathcal{N}$ is a random vector sampling a zero-mean unit-variance Gaussian distribution.

@@

We compare B2B to four standard methods: Forward regression, Backward regression, CCA and PLS, as implemented in scikit-learn \citep{scikit}, and optimized with nested cross validation over twenty l2 regularization parameters logarithmically spaced between $10^-4$ and $10^4$ (for regression methods) or 1 to 4 canonical components (for cross-decomposition methods).

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data.

@@

The chapter corresponding results can be found in Figure \ref{fig:meg_results}. Significant ($p<0.05$) pairwise differences across models are highlighted with the top horizontal bars in \ref{fig:meg_results}.

@@

Supplementary Figure \ref{fig:meg_supp} shows that the Backward model decodes the dummy variable well above chance level. In addition, "Word Length" and "Word Frequency" present a similar decoding time course, even though these features are known to specifically influence early and late MEG responses respectively \citep{kutas2011thirty}. These results illustrate the well-known fact that backward modelling cannot be used to estimate the causal contribution of collinear features.

We thus focus on the valid methods (i.e. Forward Regression, PLS, CCA, and B2B) and estimate the change of correlation coefficients $\Delta R$ induced by the introduction of each feature into the model \ref{algorithm:b2b_fi}. As expected, none of the models significantly predicted the 'Dummy Variable' to be a causal factors (i.e. $\Delta R < 0$ for all models, all $p > .089$)

@@

On average across time, "Word Length" and "Word Frequency" improved the prediction performance of all methods $\Delta R>0$ (all $p<0.0001$). As expected, the time course of "Word Length" and "Word Frequency" rose from $\approx$ 100 ms and from $\approx$ 400 ms respectively. Furthermore, 'Word Function' improved the prediction performance of all models (all $p < 0.0002$) except PLS ($p=0.7989$). Overall, these results confirm that "Word Length", "Word Frequency" and "Word Function" each causally influence specific periods of brain responses to words.

To assess which model would be more sensitive to these causal discoveries, we performed pairwise comparisons between models and across subjects (Figure \ref{fig:meg} right). For "Word Length" B2B outperforms all models (all $p < 0.00001$) but CCA ($p=0.0678$). For "Word Frequency", B2B outperforms all models (all $p < 0.0006$). For "Word Function", B2B outperforms all models (all $p < 0.0015$). Overall, these results suggest that B2B reliably outperform standard methods, especially when the effects are difficult to detect.

>>>>

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is displayed in Figure \ref{fig:meg_evoked}.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but colinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is expected to specifically cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and was derived with the the Zipf logarithmic scale of \citep{van2014subtlex} provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is expected to specifically cause a variation in the late evoked MEG responses (i.e. from 400 ms), because it variably engages semantic processes in the temporal cortices \citep{kutas2011thirty}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition, a conjunction, a determinant, a pronoun or a numeral), and was derived from Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has not been thouroughly investigated with MEG. Its causal contribution to reading processes in the brain thus remains unclear.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms thus have a zero-mean and a unit-variance).

@@

We compare B2B to four standard methods: Forward regression, Backward regression, CCA and PLS, as implemented in scikit-learn \citep{scikit}, and optimized with nested cross validation over twenty $l2$ regularization parameters logarithmically spaced between $10^{-4}$ and $10^4$ (for regression methods) or 1 to 4 canonical components (for cross-decomposition methods).

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data, using a 5-fold cross-validation.

@@

The corresponding results can be found in Figure \ref{fig:meg_results}.

@@

Supplementary Figure \ref{fig:meg_supp} shows that the Backward model decodes the dummy variable well above chance level. In addition, the Backward model reveal a similar decoding time course for Word Length and Word Frequency, even though these features are known to specifically influence early and late MEG responses respectively \citep{kutas2011thirty}. These results illustrate that backward modelling cannot be used to estimate the causal contribution of collinear features.

We thus focus on the four remaining methods (i.e. Forward Regression, PLS, CCA, and B2B) and estimate their $\Delta R$ (i.e. the improvement of Y prediction induced by the introduction of a given feature into the model \ref{algorithm:b2b_fi}). Contrary to the Backward Model, none of the models predicted the Dummy Variable to improve the $Y$ prediction: all $\Delta R < 0$ (all $p > .089$).

@@

Word Length and Word Frequency improved the prediction performance of all methods: $\Delta R>0$ for all models (all $p<0.0001$). As expected, the time course associated with Word Length and Word Frequency rose from $\approx$ 100 ms and from $\approx$ 400 ms respectively. Furthermore, Word Function improved the prediction performance of all models (all $p < 0.0002$) except PLS ($p=0.7989$). Overall, these results confirm that Word Length, Word Frequency and Word Function causally influence specific periods of brain responses to words.

To assess which model would be most sensitive to these causal discoveries, we compared B2B to other models across subjects (Figure \ref{fig:meg} right). For Word Length B2B outperforms all models (all $p < 0.00001$) but CCA ($p=0.0678$). For Word Frequency, B2B outperforms all models (all $p < 0.0006$). For "Word Function", B2B outperforms all models (all $p < 0.0015$). Overall, these results show that B2B reliably outperform standard methods, especially when the effects are difficult to detect.
